\section{\bf \textsf{Appendix C: Support documentation}}

\subsection*{Resignation letter (IFAE)}


\begin{quotation}
Dear Gloria and Juanjo (cc. IB board),In the last days, we have been carefully evaluating the recently approved NEXT CDR, particularly regarding IFAE participation. We have balanced pros and cons and we have concluded that the IFAE group will not sign this CDR for a 100 kg detector. There are many reasons for this but they can be condensed in four points:
\begin{enumerate}
\item the proposed technology has not been demonstrated to be a viable one, neither from the performance nor from the construction points of view.
\item We consider the proposed schedule as too challenging in the scope of the previous point and considering the experimental progress towards this design since the LoI 2 years ago. Therefore, we have serious doubts that the experiment will be taking data in 2014 as a competitive alternative to EXO or any other of the main competitors in the DB field on the 100 kg scale.
\item In our opinion, the technology election procedure has not been appropriate, given the lack of a consensus within the collaboration or the evaluation from and independent, external review panel.
\item The lack of information and scientific discussion within the collaboration, particularly during the critical past months, has led us to lose confidence in the capabilities of the collaboration’s management bodies to coordinate and integrate the efforts of the NEXT groups towards the important challenges the collaboration will face in the near future.Given all this, we hereby waive our membership in the NEXT Collaboration, wishing you sincerely all the best with this challenging experiment.
\end{enumerate}
 Javier Rico and Federico Sanchez
\end{quotation}

\subsection*{Report of ANEP referee to the 2011 panel (extract)}


\begin{quotation}
\item El IP tiene un historial lleno de luces y de sombras, y sus logros cient\'ificos no est\'an tan claros como \'el expresa\footnote{A subjective ``gut feeling'' statement, not backed by fats}. \'Ultimamente dedica m\'as tiempo a la escritura de novelas u otros libros que a ocuparse de los experimentos y de los consolider en los que tiene muchas responsabilidades\footnote{Ad hominem}.
\end{quotation}