
\documentclass[a4paper,11pt,oneside]{article}
\usepackage[a4paper,vmargin={1.5cm,1.5cm},width=16cm]{geometry}
%\usepackage[style=verbose-inote,doi=false,sortcites=true,block=space, backend=bibtex]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage[spanish]{babel}
\usepackage{microtype}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{eurosym}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
%% Added to help mimic structure.
%\usepackage[many]{tcolorbox}
\usepackage{tcolorbox}
\usepackage{soul}
\usepackage{color}
\usepackage{lastpage}

%\documentclass[11pt,a4paper]{article}
%\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}
%\usepackage[a4paper,height=26cm,width=17cm]{geometry}
%\usepackage{graphicx} d
%\usepackage{epsfig,rotating}
%\usepackage{amssymb}
%\usepackage{mathrsfs}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{multirow}
%\usepackage{eurosym}
%\usepackage{dcolumn}
%\usepackage{cite}

\usepackage{multirow}
%\usepackage{enumitem}

%\usepackage[hyphenbreaks]{breakurl}
%\usepackage[hyphens]{url}

\usepackage{hyperref}
\hypersetup{breaklinks=true}


%\usepackage[hyphens]{url}



%% Hack to make math formulas bold in section titles
\makeatletter
\DeclareRobustCommand*{\bfseries}{%
  \not@math@alphabet\bfseries\mathbf
  \fontseries\bfdefault\selectfont
  \boldmath
}
\makeatother

\begin{document}

\input{../Commands.tex}

\begin{center}
{\LARGE \bf \textsf{CONSOLIDER-INGENIO 2010 PROGRAMME }} \\ \vspace{0.5cm}
{\Large \bf \textsf{FINAL SCIENTIFIC ANNUAL REPORT}} \\ \vspace{0.5cm}
{\Large \bf \textsf{I01/01/2012 al 31/12/2012 }} \\ \vspace{2cm}
\end{center}


\begin{table}[htp!]
\begin{center}
\begin{tabular}{|c|c|}
\hline
PROJECT REFERENCE NUMBER: & CDS2008-00037\\
\hline
Coordinating Researcher:  & M.C. G\'onzalez-Garc\'ia \\
Project Title: & Canfranc Underground Physics (CUP) \\
Managing Institution:  & CSIC \\
Project Start Date:  & 15/12/2008\\
Project Final Date: &  14/12/2014)\\
\hline\hline
\end{tabular}
\end{center}
\caption{Project data}
\label{default}
\end{table}%

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 1
\section{\bf \textsf{CUP: summary and main goals }}

CUP (Canfranc Underground Physics) was granted with the main goal of developing the scientific program of the
 \emph{Laboratorio Subterr\'aneo de Canfranc} (LSC), one of the ICTS (instalaci\'on cient\'ifico t\'ecnica singular) of the ministry of science (currently integrated in the MINECO). The two main activities of CUP were called
 CAFE (Canfranc Future Experiment) and NEXT (Neutrino Experiment with a Xenon TPC).

CAFE has spawn a wide range of phenomenological studies, coordinated by professor 
M.C. Gonz\'alez-Garc\'ia. These studies have involved several institutions of the CUP consortium, including the Universidad de Barcelona (UB), Instituto de F\'isica Corpuscular, (IFIC), and Instituto de F\'isica Te\'orica (IFT). 

The goals of the subproject CAFE were:
\begin{enumerate}
\item The characterisation of LSC to demonstrate its viability 
as a large underground European laboratory in the context of
the European project LAGUNA (Large Apparatus for Studying Grand
Unification and Particle Astrophysics) as well as in the context
of studies of viability for future neutrino factories which were
performed as part of the European project EURO-$\nu$.
\item Perform phenomenological studies to explore and exploit 
the physics potential LSC  as well more broad studies related
to astroparticle physics in general (neutrino physics, cosmology, etc.)
\end{enumerate}

%The CAFE activities were amply described in our previous reports, and here we will simply update the lists of publications related with the activity. 

%Two major spinoffs of the activity are two European projects directly related with CAFE goals: 
%
%\begin{enumerate}
%\item 
%A european training network, called INVISIBLES, coordinates by professor B. Gavela, from IFT. 
%%
%\item A european project of large infrastructures called LAGNA-LBNO. The spanish IP of the project is  professor J.J. Gómez-Cadenas, for IFIC, who acts also as spokesperson of NEXT.
%\end{enumerate}

CUP main goal was the construction, commissioning and operation of the NEXT detector\footnote{\url{http://next.ific.uv.es/next/}}, a high-pressure, xenon (HPXe) Time Projection Chamber (TPC), whose goal was to search for neutrinoless double beta decay  (\bbonu) events in xenon enriched at 90\% in the isotope \XE. The first phase of the experiment, called NEW, deploying 10 kg of xenon, is currently being commissioned at the LSC.The second phase of the experiment, NEXT-100, will deploy 100 kg of xenon. A third phase, deploying up to one ton of xenon, is actively being discussed, with strong international interest.
 
%
%The discovery potential of NEXT is very large. It combines four desirable features that make it an almost-ideal experiment for \bbonu\ searches, namely: 
%\begin{enumerate}
%\item Excellent energy resolution (better than 1\% FWHM in the region of interest).
%\item A topological signature (the observation of the tracks of the two electrons).
%\item A fully active, very radiopure apparatus of large mass.  
%\item The capability of extending the technology to much larger masses.
%\end{enumerate}
%
%The project has evolved very satisfactorily, from the initial Letter of Intent (LOI) in 2009 to the Technical Design Report (TDR) in 2012. A substantial number of papers, proceeding reports and conferences, documenting and demonstrating the physics case, the results of the prototypes and the technological choices have been published and are attached to this document. They can also be found in the NEXT web page: \url{http://next.ific.uv.es/next/talks.html}.
%


\section{\bf \textsf{Main advances and profits }}
\label{sec.advances}
\input{src/AdvancesAndProfits.tex}

\section{\bf \textsf{ Key Indicators}}
\label{sec.indicators}

\subsection*{Reports from the LSC Scientific Committee}

In the CUP proposal, we stated:
\begin{quotation}
We propose a very simple and efficient evaluation scheme for CUP. The  activities are focused in the LSC, which has an international scientific committee, composed by high-reputation specialists in underground physics. This committee is perfectly suited to evaluate our progress in all the areas proposed.
\end{quotation}

The reports of the LSC Scientific Committee (LSCSC) have been consistently encouraging and enthusiastic towards NEXT. They represent a continuous and methodic follow up of a very difficult and complex project, and they
also keep a detailed historical track of the steady progress of the experiment. At the same time, the reports reflect the concerns of the LSCSC regarding a perceived lack of domestic support to the project, in particular during 2011 and 2012.
It is important to remark that the LSCSC is an independent body, formed almost exclusively by experts of international reputation. This adds tremendous weight to the opinions, recommendations and commendations expressed in their reports and permits an objective evaluation of the degree of success of the project. 

{\em
In appendix A, we offer a detailed discussion of the LSC reports. 
}

\subsection*{Other key indicators for NEXT. Publications, international projection and grants}

\begin{itemize}
\item {\bf Publications}: the project has generated a large number of publications in refereed journals and conferences, as well as posters and communications. A full list of publications can be found online\footnote{\url{http://next.ific.uv.es/next/talks.html}}.
\item{\bf International projection}: NEXT is a CERN recognised experiment and is included among the most important
\bbonu\ experiments in Europe. See for example the recent presentation of A. Giuliani in the APPEC meeting \footnote{\url{http://app2016.in2p3.fr/programme.html}}.
\item{\bf Grants:} The spokesperson of NEXT has obtained and Advanced Grant of the ERC for the NEXT project. 
\end{itemize}

\subsection*{CAFE: key indicators}

%\begin{quotation}
%As for the CAFE activity, there is an obvious evaluation criteria, namely the scientific papers to be produced along the five years of the project, including the proceedings of the workshop(s) and/or schools as well as other outreach material.
%\end{quotation}
%
The key indicators of the final results of CAFE are the following:

\begin{enumerate}
\item Involvement in the project LAGUNA-LBNO
(\url{http://laguna.ethz.ch:8080/Plone}). This project continued the
studied started by LAGUNA with the aim at developing large infrastructures
in Europe. Spain participated with four associates LAGUNA-LBO, three of
which are participants in CUP (UAM, IFIC y el propio LSC).
\item Involvement of our scientists in the european project EURO-$\nu$.
In particular de group at UAM made very relevant contributions to
this project as listed in the bibliography.
\item Participation in three funded projects of the European Commission
for the training of young researchers in fields directly related
with underground physics. They are: 
\begin{itemize}
\item 
Title: INVISIBLES\\
Initial year: 2012 Year final: 2016\\
Reference: PITN-GA-2011-289442\\
\item 
Title: 	Elusives\\
Initial year: 2016 Year final: 2020\\
Reference: H2020-MSCA-ITN-2015-674896 \\
\item  Title: 	InvisiblesPlus\\
Initial year: 2016 Year final: 2020\\
Reference: H2020-MSCA-RISE-2015-690575\\
\end{itemize}
The coordinator of the UB node of these networks is our IP, Prof.
M.C. Gonzalez-Garcia while the coordinator of the node at IFIC
is Prof. P. Hernandez, a member of CAFE/CUP.
\item Publication in peer review international journal 
of more than 70 articles in phenomenological studies of which 
we list a selection in the
bibliography. They have accumulated more than 3000 citation  in the
basis SPIRES. For example our updated determinations of the neutrino 
parameters, fundamental for any study of the expected signal at NEXT
\cite{Gonzalez-Garcia:2014bfa,GonzalezGarcia:2012sz,GonzalezGarcia:2010er} 
have accumulated more than 1000 citations,  
while our contributions to studies associated to future neutrino facilities,
such as Ref.~\cite{Bandyopadhyay:2007kx,Abazajian:2012ys,Choubey:2011zzq,Adey:2014rfv} have received more than 500 references (all data from SPIRES
Data basis).
\end{enumerate}

%

\section{\bf \textsf{ Difficulties}} 
\label{sec.diff}
The main difficulty the project has faced was related with the withdrawal, in 2011, of the IFAE and CIEMAT groups from the NEXT collaboration (although they did not formally leave the CUP consortium). The reasons argued for the withdrawal were related with different point of views concerning the technical choices implemented in NEXT, as well as disagreement with the leadership of the NEXT spokesperson.

A few months after the withdrawal of the IFAE and CIEMAT groups, the regular research project submitted by IFIC-UPV was marked with at C (mediocre) and rejected by the funding committee of particle physics (FPA). The external expert of that particular committee was prof. Giorgio Gratta (the spokesperson of the EXO experiment, which is the main competitor of NEXT) and the committee members included several leaders of the IFAE and CIEMAT group. The committee disregarded the reports of the LSC scientific committee and the opinion of the director of the LSC. The reports from the ANEP were requested and at least one of them was found to be extremely biased, with ad-hominem attacks and absolute lack of rigor, as acknowledged by the ANEP itself. However, the committee of particle physics did not reject that particular report.

The project submitted in 2011, requested essential additional funding, not covered by CUP. The negative result implied a major logistic drawback and a serious blow for the collaboration. In 2012, a new research project was submitted, to a different committee (general physics, FIS), given the obvious bias of the FPA committee. The project obtained a B and modest funding. In 2013, the spokesperson of NEXT was granted the first (and the only one, to date) advanced grant of the ERC in experimental particle physics in Spain. In 2014, another project was submitted to FIS, which obtained a top mark. 

The perception of the consortium and the international collaboration is that, in spite of the demonstrated excellence of the project (availed by the reports of the LSC scientific committee as well as by the excellent reports produced by external evaluators, in particular in the ERC), its capability to attract external funding, and its importance for the scientific case of the LSC, a bias still exist in FPA, mostly related with the opposition to the project of the IFAE group. Indeed, the ex-director of IFAE, Dr. Mateo Cavalli-Sforza has recently (in 2015) expressed publicly (in a seminar at the ALBA seminar) his enmity towards the spokesperson of the experiment. 

The situation is not fully solved yet, and poses a serious threat to the future of the collaboration, since external collaborators from other countries may feel that the project is not sufficiently supported in Spain. It also poses a serious threat for the future of the Canfranc laboratory, whose reputation would suffer very much should the NEXT experiment be forced to move to a different underground laboratory. 

The rejection of the research proposal in 2011 and the very scarce funding in 2012, forced the collaboration to redefine its initial plan to start NEXT-100 construction in 2012. Instead, it was decided to introduce a first-stage experiment, NEW (a detector with half of the size of NEXT-100, a fourth of the sensors and 10 kg of xenon mass). It was perceived that there was not sufficient support in Spain to attempt the construction of NEXT-100 directly. After the ERC grant the situation has improved considerably, but the collaboration decided to continue with the NEW program. Currently, NEW is starting data taking at the LSC, and the lessons learned in its construction (and certainly by operating it) are very valuable to build a better NEXT-100 detector. 


\section{\bf \textsf{ Scientific and technical activities }} 

\subsection*{NEXT: Scientific and technical progress}

The NEXT experiment has evolved, largely thanks to the support of the Consolider CUP, from a visionary idea ({\em to setup an international experiment at the LSC able to be in the forefront of the Underground Physics field, with a high discovery potential}) to an established reality. After an initial period of R\&D (2010 and the first half of 2011), the project has evolved in a consistent way, first with the construction, commissioning and operation of NEXT-DEMO (2011--2013), which has produced a large scientific and technical output, then with NEW which started in earnest in 2014 and will carry its scientific program through 2018. The current plan is to present an update TDR in November of 2016, and start the construction of NEXT-100 in mid/late 2017 or 2018, depending on the available human resources and funding. 
 
 NEXT-DEMO and NEXT-DBDM were essential to fully demonstrate the performance of the HPXe-EL technology, including:
 \begin{enumerate}
\item An energy resolution near the Fano-Factor ($\sim$ 0.5 \% FWHM at \Qbb)\cite{Alvarez:2012yxw,Alvarez:2012zsz}.
\item First operation in the World of a tracking plane made of SiPMs \cite{Alvarez:2013gxa}.
\item Demonstration of the unique topological signature available in HPXe\cite{Ferrario:2015kta}.
\end{enumerate}

During the NEW phase of the project, the collaboration has addressed both scientific and technical challenges. In the scientific front:

\begin{enumerate}
\item Understanding of the radioactive budget \cite{Cebrian:2015jna, Alvarez:2014kvs, Dafni:2014yja}.
\item Definition of the detector sensitivity \cite{Martin-Albo:2015rhw}.
\item Demonstration of the unique topological signature available in HPXe\cite{Ferrario:2015kta}.
\end{enumerate}

And in the technical front:

\begin{enumerate}
\item Construction and certification of a large underground detector (NEW) for underground operation.
\item Construction and certification of large infrastructures (gas system, high-voltage system) for for underground operation.
\item Development of slow controls.
\item Risk analysis.
\item System integration.
\end{enumerate}

With the acquired experience, the collaboration is looking forward to an steady operation of NEW followed by the
timely construction, commissioning and operation of the NEXT-100 detector. The NEXT program is one of the major alternatives in Europe for a future, ton-scale, \bbonu\ experiment. 


\subsection*{CAFE: Scientific activity}

The scientific activity of CAFE has been described above. We list
here the main lines developed in their phenomenological studies
which are directly related to underground physics
\begin{itemize}
\item Precise and updated characterization of the leptonic flavour
parameters, whose values are crucial to the prediction of expected
event rates in neutrinoless double beta decay experiments
\cite{Song:2015xaa,Bergstrom:2015rba,Gonzalez-Garcia:2014bfa,Gonzalez-Garcia:2013usa,GonzalezGarcia:2012sz,GonzalezGarcia:2011my,GonzalezGarcia:2010un,GonzalezGarcia:2010er}
\item Studies of new phenomena expected from presence of additional
sterile neutrinos and their effect at future underground experiments
\cite{Vincent:2014rja,Bergstrom:2014fqa,Adey:2014rfv,GonzalezGarcia:2012yq,SolagurenBeascoa:2012cz,Donini:2011jh,Abazajian:2012ys}
\item Studies directly related to future underground facilities
in Europe 
\cite{Bross:2013oua,Edgecock:2013lga,Blennow:2013swa,Bayes:2012ex,Agarwalla:2012zu,Hernandez:2012zz,Coloma:2012wq,Coloma:2011rq,Donini:2010xk,Coloma:2010wa,Choubey:2009ks,Bandyopadhyay:2007kx}
\item Potential of underground experiments to detect and 
characterize astrophysical neutrino sources 
\cite{Bergstrom:2016cbh,Gonzalez-Garcia:2013iha,Ahlers:2011jj,Ahlers:2010fw,GonzalezGarcia:2009ya,GonzalezGarcia:2009jc}
\item Study of theoretical frameworks for the connection  between Majorana 
nature of the neutrino (as searched for
in neutrinoless doublebeta decay) and the production of the
observed matter-antimater in the Universe
\cite{Hernandez:2015wna,Fong:2013gaa,Fong:2011yx,Fong:2010bv,Fong:2010bh,Fong:2010qh,Fong:2010zu,GonzalezGarcia:2009qd,Fong:2009iu}
\item Construction of models which can account for all neutrino observations
and study their specific predictions in future experiments
(including neutrinoless doublebeta decay experiments)
\cite{Gago:2015vma,Hernandez:2014fha,Hernandez:2013lza,Donini:2012tt,Eboli:2011ia,Gavela:2009cd}
\item Studies of scenarios of electroweak symmetry breaking related
for mass generation
\cite{Corbett:2015lfa,Corbett:2014ora,Gavela:2014vra,Brivio:2014pfa,Brivio:2013pma,Corbett:2013pja,Corbett:2012ja,Corbett:2012dm,Eboli:2011ye,Eboli:2011bq,Eboli:2010qd,Alves:2009aa}
\end{itemize}
%


\section{\bf \textsf{Collaborations between team members}}

\subsection*{NEXT}

The scientific and technical activities of NEXT are carried out in the context of an international collaboration which includes the members of the CUP consortium but also other international members, such as the groups from Portugal, Colombia and the US. 

The leadership of the collaboration rests largely in Spanish groups, and in particular in the leading groups in the CUP consortium. Specifically:

\begin{itemize}
\item The spokesperson (J.J. Gomez-Cadenas) is a member of the IFIC group.
\item The hardware project manager (J.F. Toledo) is a member of the UPV.
\item The software project manager (J.A. Hernando) is a member of the U. Santiago (US).
\item The technical coordinator (F. Monrabal) is a member of U. of Texas at Arlington and a former Ph.D. student of NEXT.
\item The analysis convener (M. Sorel) is a member of IFIC.
\item The software coordinator (P. Novella) is a member of the IFIC. 
\item The radiopurity coordinator (S. Cebrian) is a member of U. Zaragoza.
\item The integration manager (J. Torrent) is a member of U. Zaragoza.
\item The DAQ coordinator and the Slow Control coordinators are member of the UPV.
\item The project leaders of detector construction are members of IFIC.
\end{itemize}


\subsection*{CAFE: Scientific activity}

As can be seen from our publication list there are regular
collaborations between the scientists in UB, IFIC and UAM. 
The members in the three groups in CAFE are all involved in the European
networks mentioned listed above. 


\section{\bf \textsf{Collaborations with other groups of research}}

7.-Relate the collaborations with other groups of research and the added value for the project. Describe, if necessary, the access to equipment or infrastructures of other groups or institutions

\subsection*{NEXT}

The international part of the NEXT collaboration has played a very important role in the dynamics of the experiment. In particular, the impact of the US groups cannot be overemphasised. Prof. D. Nygren was the inventor, in the mid seventies, of the TPC technology which constitutes the backbone of both the NEXT and EXO experiment (together with many other experiments in underground physics, nuclear physics and high energy physics). Nygren is a recipient of numerous prices and awards, a long-time distinguished scientist at Berkeley (now holding a chair at U. of Texas at Arlington) and one of the most influential scientist of his generation. The late prof. J. White was a World expert in high-pressure xenon chambers, and his expertise was essential to build the NEXT-DEMO detector.

The Spanish group at IFIC, UPV and elsewhere, benefited enormously from the experience, know-how and ideas of this two leading scientist and their teams. Furthermore, NEXT has been collaborating with scientists from Fermilab national laboratory. NEXT has been presented to the Nuclear Science Committee (NSAC) and has deserved excellent review
\footnote{\url{http://science.energy.gov/np/nsac/
}}

The US groups have contributed also importantly to the funding of the experiment, with man power and contributions in cash and equipment, currently at the level of $\sim$300 k\euro. 

The Portuguese groups have also contributed with their know-how (the groups of Coimbra and Aveiro are recognized experts in the field) and have lead several important studies in connection with energy resolution and gas mixtures. The Russian group from Dubna was essential for the procuration of the 100 kg of enriched xenon owned by the LSC and necessary for the experiment (the current cost of enriched xenon is five time the cost payed in 2011). Last, but not least, the group from A. Nari\~no (Colombia) is contributing in several aspects of software and DAQ. 

{\em The success of NEW and a clear message of support from the Spanish authorities would be instrumental in enlarging the NEXT collaboration}. The construction and full commissioning of the NEXT-100 detector could benefit enormously from a stronger international collaboration, and the costs could be shared among more partners. 

\subsection*{CAFE}

We have ongoing collaborations with scientist in major institutions
and laboratories worldwide as can be deduced from our list
of publications. Just to mention a few: CERN, Fermilab (USA), 
INFN-Frascati (ITALY), Gran Sasso (Italy), Max Planck Institute (Germany),
Stony Brook Univ (USA), U. Wisconsin (USA), and, U. Sao Paulo (Brazil). 

\section{\bf \textsf{Collaborations with companies}}

The NEXT project has relations with a large number of companies. Among these:
\begin{itemize}
\item Hamamatsu Photonics and SENSL, for sensor developments (radiopure PMTs, radiopure SiPMs). We have an on-going collaboration with both companies, in particular to develop ultra-pure SiPMs.
\item Linde, Air Liquide and Swagelock, concerning gas system developments. 
\item ACYM and other pressure vessel manufacturing companies. 
\item SCIENTIFICA and AVS, which are science-related companies with whom we have also collaboration agreements. 
\end{itemize}


\section{\bf \textsf{UE FP7 calls}}

\begin{itemize}
\item 
Title: INVISIBLES\\
Initial year: 2012 Year final: 2016\\
Reference: PITN-GA-2011-289442\\
\item 
Title: 	Elusives\\
Initial year: 2016 Year final: 2020\\
Reference: H2020-MSCA-ITN-2015-674896 \\
\item  Title: 	InvisiblesPlus\\
Initial year: 2016 Year final: 2020\\
Reference: H2020-MSCA-RISE-2015-690575\\
\item Title: NEXT \\
Initial year: 2014 Year final: 2019\\
Reference: FP7-Advanced Grant--339787.
\end{itemize}

\section{\bf \textsf{National and international Program visibility}}

10.-National and international Program visibility

CAFE: The members of CAFE been regularly reporting the results of their 
studies in  conferences and workshops. In particular they have been invited to
be review talks on neutrino physics in the most important conference
in the area. For example:
\begin{itemize}
\item Prof. M.C. Gonzalez-Garcia gave the plenary talk on Theory of Neutrino
Physics in the ICHEP Conference in Melbourne in July 2012
\item Prof. P. Hernandez gave the plenary talk on Theory of Neutrino Physics
at the HEP-EPS Conference in Vienna in July 2015 
\item Prof. M.C. Gonzalez-Garcia gave the plenary talk on Neutrino Physics
at the  Astroparticle Physics Joint  TeVPA/IDM Conference, Amsterdam, 
Netherlands in June 2014
\end{itemize}

NEXT presentes regularly its status and results in international conferences, see
\url{http://next.ific.uv.es/next/talks.html}. Prof. Gomez-Cadenas has given numerous international talks and seminars as well as
taught in many international schools. For example, in 2015:

\begin{itemize}
\item Gave a plenary talk on Experimental Neutrino Physics
at the HEP-EPS Conference in Vienna in July 2015.
\item Was convener of the \bbonu\ discussion group at the Aspen Physics Center, August, 2015.
\item Gave an invited plenary talk at the Colloquium Spectroscopicum Internationale (CSI), Septembre 2015.
\end{itemize}

Other talks and seminars:

\begin{itemize}
\item \textit{Ettore Majorana through the looking glass (searching for neutrinoless double beta decay)}, Harvard Monday Colloquium, 22 October 2012; Fermilab Colloquium, 24 October, 2012; University of Wisconsin Madison seminar, 26 October 2012.
%% 2
\item \textit{NEXT, high-pressure xenon gas experiments for ultimate sensitivity to Majorana neutrinos}, invited talk at the 14th International Workshop on Radiation Imaging Detectors (iWoRID 2012), Figueira da Foz, Coimbra (Portugal), 1-5 July, 2012. 
%% 3
\item \textit{Xenon for DM and DBD searches}, invited talk at IDPASC Dark Matter Workshop, \'Evora (Portugal), 2011.
%% 4
\item \textit{Status of the NEXT experiment}, at DBD'11: International Workshop on Double Beta Decay and Neutrinos, Osaka (Japan), 2011.
%% 5
\item \textit{How to probe anti-neutrino = neutrino and the absolute neutrino mass scale}, International Neutrino Summer School, Geneva (Switzerland), 2011.
%% 6
\item \textit{Sense and sensitivity in \bbonu\ experiments} at XIV International Workshop on Neutrino Telescopes, Venice (Italy), 2010.
%% 7
\item {\it Ettore Majorana meets his shadow (searching for neutrino less double beta decay)}, Wednesday colloquium, Weizmann institute, 24 November 2010.
\end{itemize}


\section{\bf \textsf{Problems and suggestions}}

The Consolider CUP, together with the AdG from the ERC and the support of the SEIDI has made possible to launch the NEXT experiment. NEXT is the flagship experiment at the LSC, as clearly recognised by the LSCSC (see appendix), and its success will be a tremendous boost for the credibility and visibility of our laboratory. NEXT has a large discovery potential and is considered one of the most relevant technologies for the future of t\bbonu\ searches. Very likely there will be a ton-scale HPXe experiment searching for \bbonu\ events in the next decade. Having launched the NEXT program, Spain is in a perfect position to host such a large-scale experiment, involving a large international collaboration at the LSC, leading the field. 

However, the project still faces difficulties in Spain. A large experiment like NEXT requires a stable funding scheme like the CONSOLIDER program which is, alas, extinct. CUP made possible the creation of groups of young scientists and engineers at IFIC, UPV, U. Santiago and U. Zaragoza, among others. In particular the IFIC group includes about 5 post-docs and engineers who play leading roles in the development of the experiment. None of these people have permanent positions. Their salaries are payed either by a SEIDI grant or by the ERC grant. In the medium term this situation is not sustainable. If the IFIC group dismembers due to lack of funds or to lack of long-term prospects for the physicists and engineers working there, the know-how and capability to lead will be lost and the experimental push at the LSC will stale. Another opportunity will be lost. 

Unlike most efforts in particle physics, where Spain contributes with sizeable funds to experiments happening elsewhere (and more often than not, lead by non-Spanish physicists), NEXT is an example of an international experiment being led by Spanish physicists and conducted in Spain. NEXT has also attracted considerable external funding and prestige, as is the case also with CAFE. The two projects together participate or have participated in 6 EC projects (LAGUNA-LBO, Euro-$\nu$, INVISIBLES, INVISIBLES-PLUS, ELUSIVES and the AdG). However, NEXT has been consistently under-funded compared with other projects in the particle physics community, as already discussed. 

We hope that the situation can be improved.  

\section{\bf \textsf{Appendix A: Reports of the LSC Scientific Committee }}
In this section we present the reports of the LSC Scientific Committee (LSCSC) since 2010, together with a discussion concerning the implementations of the committee recommendations. The committee reports are copied verbatim. {\em Emphasis} is used to highlight key aspects of the text, while {\bf bold face} is only used if the committee used it in the original report. Footnotes, often added to specific recommendations, add relevant comments or reference to bibliography. 

\input{lscsc/2010.tex}
\input{lscsc/2011.tex}
\input{lscsc/2012.tex}
\input{lscsc/2013.tex}
\input{lscsc/2014.tex}
\input{lscsc/2015.tex}

%\bibliographystyle{elsarticle-num}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}

